---
title: "Masters work"
author: "Elizabeth Okello-113285"
date: "2023-03-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Lets load the dependencies
```{r}
library(ggplot2)#for visualization
library(forecast)#for forecasting
library(dplyr)#for data manipulation
library(tseries)#for timeseries analysis
library(rugarch)#for ARCH/GARCH models
library(e1071)
library(FinTS)#to check for ARCH effects
#require(forecast)
#install.packages("PerformanceAnalytics")
library(xts)
library(quantmod)
library(PerformanceAnalytics)
library(vars)#for using VARS
library(lmtest)#for bptest
library(broom)#for table presentations
library(zoo)
library(strucchange)
library(moments)
library(astsa)
library(tidyverse)
```


#Lets load the data
```{r}
profercy<-read.csv("C:/Users/spectre/OneDrive/Documents/time series content/profercy.csv",header=T,skip=3)
profercy[,c("Week.No.","X")]<-list(NULL)#to drop the empty columns
head(profercy)
```
#Lets get the structure of the data and check for missing values
```{r}
str(profercy)#to check the structure of the data
sum(is.na(profercy))#to check for missing values
```
The data has 1254  missing values

#Lets remove the missing values
```{r}
profercy.complete<-na.omit(profercy)#to remove missing values

sum(is.na(profercy.complete))
#View(profercy.complete)
```
#We have two ports of interest so lets subset each of them
#our working variables will be the month.year and the port of Yuzhny so lets subset them
```{r}
yuz<-profercy.complete[,c("Month.Year","P.UREA.Yuzhny.FOB")]
#View(yuz)
str(yuz)
```
#Lets write it into a ts objects and visualise
#port of  YUZ.
```{r}
yuzts<-ts(yuz[,2],start=c(2004,1),end=2018,frequency = 12)#TO write into a ts object

plot(yuzts,xlab="Years",ylab="price")#to plot the ts object
ggtsdisplay(yuzts)

```
From here we see a clear indication of non-stationarity in the ACF and a strong partial autocorrelation for the first two lags
#Lets decompose the data for YUZ into its components
```{r}
dec<-decompose(yuzts)
plot(dec)
```
From the decomposed time series seasonality and trend are present

#Lets plot the acf and pacf before transforming the data
```{r}
acf(yuzts)#To plot acf 
pacf(yuzts)#To plot pacf
```

This confirms that the data is non stationary
```{r}
print(kurtosis(yuzts))
print(skewness(yuzts))
sd(yuzts)
```
There is a mild skewness

#Lets test for normalility
```{r}
shapiro.test(yuzts)#To test for normality
```
The p-value is less than 0.05 ie (p-value = 2.176e-05) and therefore we reject the null hypothesis and conclude that the distribution is significantly different from a normal distribution(not normally distributed)

#Lets plot the histogram to check
```{r}
hist(yuzts)#To plot histogram
```
This reinforces our first result though not very far from normal. We may not transform it.

#Lets test for stationarity for yuzts
```{r}
adf.test(yuzts, k=2)#we use the adf test to check for stationarity
```
From the above results the p-value(0.325) is greater than 0.05 we therefore fail to reject our null hypothesis and conclude that the data is non stationary.
#Lets try a kpss test
```{r}
kpss.test(yuzts)
```
We reject the Null hypothesis that the data is stationary and conclude that the data is non stationary

#Since seasonality is present lets take the seasonal diff
```{r}
yuz_diff12 = diff(yuzts,12)#difference to make it stationary and assign to variable name 
par(mfrow=c(1,2))
acf(yuz_diff12, 48)#To plot the acf of seasonal differenced data
pacf(yuz_diff12, 48)#To plot the pacf of seasonal differenced data
```

#Lets evaluate for trend in the differenced data by decomposing the data
```{r}
dec2<-decompose(yuz_diff12)
plot(dec2)
```

#Trend is still present so lets detrend the data as well
```{r}
diff1and12 = diff(yuz_diff12, 1) #To detrend the data
par(mfrow=c(1,1))
acf(diff1and12, 48)
pacf(diff1and12, 48)
plot(diff1and12)
```

#Lets check for stationarity first for the differenced data
```{r}
adf.test(diff1and12)
kpss.test(diff1and12)
pp.test(diff1and12)
```
All the tests show that the differenced data for both the seasonality and trend is now stationary

#Lets split the data into to train and test set
```{r}
myts.train <- window(diff1and12, end=c(2016,12))
myts.test <- window(diff1and12, start=2017)
```
#Lets check how the splitting has been done
```{r}
autoplot(diff1and12) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

#Lets select the model using auto sarima on the whole data(before splitting)
```{r}
yuz_auto<-auto.arima(diff1and12,
                 stepwise = F,
                 approximation = F,
                seasonal = T,
                 allowdrift = F
                 )
#summary(yuz_auto)
```

#For the train data yuz

```{r}
yuz1_auto<-auto.arima(myts.train,
                 stepwise = F,
                 approximation = F,
                 seasonal = T,
                 allowdrift = F
                 )#Use auto arima to automatically select the best model
summary(yuz1_auto)
```
The split data train yields a low AIC

#Lets use the model to forecast
```{r}
sarima.forecast<-forecast(yuz1_auto, h=12)
plot(sarima.forecast)
```

```{r}
SARIMA.accuracy<- accuracy(sarima.forecast,myts.test)
```

#Lets graph the results
```{r}
autoplot(yuzts) +
  autolayer(sarima.forecast, series = "SARIMA Forecast") +
  autolayer(myts.test, series = "Actual PRICES")
```

The SARIMA model has forecasted farely well. however from the residuals we see heteroskedasticity. So we take the residuals to the GARCH as long as they pass the ARCH test
#Letsextract sarima residuals for the train set
```{r}
yuz1_auto_residuals <- residuals(yuz1_auto)
```
#Lets test the sarima residuals for arch effects on the train set
```{r}
ArchTest(yuz1_auto_residuals)
ArchTest(yuz1_auto_residuals^2)
```
The squared residuals pass the ARCH effects test


#Lets assign the squared residuals into a variable name
```{r}
re<-yuz1_auto_residuals^2
```

#Lets split the residuals into test and train and fit a GARCH model on them
#Lets split the residuals into to train and test set
```{r}
re.train <- window(re, end=c(2015,12))
re.test <- window(re, start=2016)
```
#Lets check how the splitting has been done
```{r}
autoplot(re) +
  autolayer(re.train, series="Training") +
  autolayer(re.test, series="Test")
```


#Lets fit the garch model using the residuals
```{r}
garch_spec <- ugarchspec(variance.model = list(model = "gjrGARCH"), 
                         mean.model = list(armaOrder = c(0,0)),
                         distribution.model = "sstd") # Define GARCH specification
garch_fit <- ugarchfit(spec = garch_spec, data = re.train)
garch_fit
```

#Letsforecast
```{r}
hybrid_model <- ugarchforecast(garch_fit, n.ahead = length(myts.train))
#plot(hybrid_model)
#garch_sigma <- ugarchforecast(garch_fit, n.ahead = length(myts.train))
#garch_sigma
```



```{r}
#forecast_results <- forecast(hybrid_model, h = 12)
#forecast_result
```






```{r}


# Evaluate model performance
forecast_values <- as.numeric(hybrid_model@forecast$seriesFor)
accuracy <- accuracy(forecast_values, myts.test) # Use accuracy function to calculate forecast accuracy
print(accuracy) # Print forecast accuracy
```


```{r}
# Make forecasts
forecast_values <- as.numeric(hybrid_model@forecast$sigmaFor) # Extract forecasted values
forecast_values

```
```{r}
# Visualize forecast
# Convert forecasted values to data frame
forecast_df <- data.frame(Date = time(myts.train), Forecast = forecast_values)
# Plot forecast
ggplot() +
  geom_line(data = data.frame(Date = time(myts.train), Value = forecast_values), aes(x = Date, y = Value), color = "black", linetype = "solid") +
  geom_line(data = forecast_df, aes(x = Date, y = Forecast), color = "blue", linetype = "dashed") +
  labs(title = "Hybrid SARIMA-GARCH Forecast",
       x = "Date",
       y = "Value") +
  theme_minimal()
```


```{r}
autoplot(residuals(yuz1_auto)) # plot the residuals 
plot(yuz1_auto) # inspect the roots  
ggAcf(residuals(yuz1_auto),lag.max=36) # inspect the ACF 
ggPacf(residuals(yuz1_auto),lag.max=36) # inspect the PACF 
Box.test(residuals(yuz1_auto), lag=24, type="Ljung-Box")
```


```{r}
autoplot(residuals(yuz_auto)) # plot the residuals 
plot(yuz_auto) # inspect the roots  
ggAcf(residuals(yuz_auto),lag.max=36) # inspect the ACF 
ggPacf(residuals(yuz_auto),lag.max=36) # inspect the PACF 
Box.test(residuals(yuz_auto), lag=24, type="Ljung-Box")
```
The Ljung-Box test uses the following hypotheses:

H0: The residuals are independently distributed.

HA: The residuals are not independently distributed; they exhibit serial correlation.


From the Box-Ljung test the p-value(0.2094) is greater than 0.05 and therefore fail to reject the null hypothesis and conclude that the residuals are independent


#Lets use the model to forecast
```{r}
fcst<-forecast(yuz_auto, n.ahead=12)
plot(fcst)
```



#Lets fit the first model
```{r}
fit1_yuz<-sarima(yuzts, 1,1,0,2,1,0,S=1,details = T,Model = T)
fit1_yuz
#fc <- snaive(myts.train)
#accuracy(fc,myts.test)
```
#Interpretations of the Diagnostics
The time series plot of the standardized residuals mostly indicates that there's no trend in the residuals, no outliers, and in general, no changing variance across time.

The ACF of the residuals shows some significant autocorrelations as there are significant spikes at various lags (lags 8, 9)Not a very good result

The Q-Q plot is a normal probability plot. It doesn't look too bad, so the assumption of normally distributed residuals looks okay.

The bottom plot gives p-values for the Ljung-Box-Pierce statistics for each lag up to 20 lags. Several values of the Ljung-Box-Pierce p-values are below .05. We don't want them there. So, the Seasonal(1,1,0,2,0,0,12) is a good model but may not be the best. We make this better by fitting a GARCH model on the residuals.



#Lets check residuals of the forecast
```{r}
checkresiduals(fcst)
```





#Lets Make Forecasts
```{r}
pred_model<- arima(yuzts, order=c(1,1,0), seasonal = c(2,1,0))# The sarima model
#themodel = arima(flow, order = c(1,0,0), seasonal = list(order = c(0,1,1), period = 12))

predict(pred_model, n.ahead=24)
plot(pred_model$residuals)
#resid(pred_model)

```

#Lets fit a garch. Lets assign the residuals to a variable name and write them to a ts object
```{r}
yts<-ts(resid(pred_model)^2)#write to a ts object and assign to a variable name
ytss<-as.numeric(resid(pred_model)^2)
YTs<-ts(ytss)
s<-rep(1:169)
s
final<-cbind(s,YTs)
View(final)
#View(final)
plot.ts(YTs)
hist(YTs)
hist(resid(pred_model))
adf.test(YTs)
#plot(abs(YTs))
qqnorm(YTs, main = "Squared resids -QQ Plot", col = "blue")
qqline(YTs)
View(YTs)
```
#The squared residuals are stationary as we reject the null hyp and conclude that the data is stationary
#From the histogram we conclude that the squared residuals are non normal
```{r}
shapiro.test(YTs)
```
The p-value is less than 0.05 ie (p-value = 2.2e-16) and therefore we reject the null hypothesis and conclude that the distribution is significantly different from a normal distribution(not normally distributed)

```{r}
#library(forecast)
#pred_vals<- forecast(pred_model, level=c(95), h=24)
#plot(pred_vals)
#pred_vals
#plot(fitted(pred_vals),pred_vals)
```

#Lets plot the acf and pacf of the squared residuals
```{r}
par(mfrow=c(1,2))
acf(YTs)
pacf(YTs)
```


#Lets test for ARCH effect son the squared residuals of YUZ

#Arch test of model residuals

```{r}

ArchTest(resid(pred_model))#Test for arch effect on residuals
ArchTest(YTs)#Test for arch effects on the squared residuals
```

Arch effects are present in the squared residuals but not in the residuals.
We therefore proceed to to fit a GARCH model on the residuals.

# GARCH Model specification
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="eGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
yfit <- ugarchfit(spec, data=re.train,solver = "hybrid")
show(yfit)
myforecast<-ugarchforecast(yfit, n.head=50)
#Lets graph this result
Plot.Garch = ts(fitted(myforecast), start = c(2014), frequency = 12)
#re.train <- window(re, end=c(2015,12))
#re.test <- window(re, start=2016)
autoplot(re) +
  autolayer(Plot.Garch, series = "Garch Forecast") +
  autolayer(re.test, series = "Actual price")
#show(myfit)

```
##lets extract the residuals of the fitted model
```{r}
#yres = residuals(yfit)
#acf(yres)
#pacf(yres)
```

#Volatility forecast
```{r}
#f<-ugarchforecast(fitORspec=yfit,n.head=24)
#f
#plot(fitted(f))
#plot(sigma(f))

```



```{r}
#plot(YTs$yuzforecast)
# Plot forecasts and test set
#autoplot(yuzforecast) + autolayer(a2)
#plot(yuzforecast)
```
#Lets adjust the distribution to a normal distrib but use the same model
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="eGARCH",garchOrder= c(1,1)),
                  distribution.model = "snorm",)
yfit1 <- ugarchfit(spec, data=re.train,solver = "hybrid")
show(yfit1)
```

#
```{r}
#plot(yfit1,which="all")

```

#Lets find other available  gjrGARCH model using sstd distribution
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="gjrGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
yfit_2 <- ugarchfit(spec, re.train,solver = "hybrid")
show(yfit_2)
```
#Lets find other available  gjrGARCH model using normal distribution
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="gjrGARCH",garchOrder= c(1,1)),
                  distribution.model = "snorm",)
yfit_3 <- ugarchfit(spec, YTs,solver = "hybrid")
show(yfit_3)

```

#Lets find other available  sGARCH models
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="sGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
yfit_4 <- ugarchfit(spec, re.train,solver = "hybrid")
yfit_4
#summary(yfit_4)
#cond_mean <- as.numeric(fitted(yfit_4))
#cond_var <- as.numeric(sigma(yfit_4)^2)
```




```{r}
# Calculate predictions for testing set
yfit4forecast<-ugarchforecast(yfit_4, n.ahead = length(re.test))
yfit4forecast

# Calculate AIC
#aic<-uGARCHfit(AIC(yfit_4))

# Calculate RMSE
rmse <- sqrt(mean((yfit4forecast$mean - re.test)^2))

# Calculate MAPE
mape <- mean(abs((yfit4forecast$mean - re.test)/re.test)) * 100

# Calculate MAE
mae <- mean(abs(garch_predict$mean - test))
```






#Lets diagnose our model

```{r}
#residuals(yfit_4)
yfit4res<-residuals(yfit_4)
acf(yfit4res)
acf(yfit4res^2)
```
```{r}
plot(yfit_4,which="all")
```

```{r}

data_plot <- data.frame(date = index(s), 
                        original_data = coredata(YTs), 
                        conditional_mean = cond_mean, 
                        conditional_var = cond_var)

ggplot(data_plot, aes(x = date)) +
  geom_line(aes(y = original_data), color = "blue") +
  geom_line(aes(y = conditional_mean), color = "red") +
  geom_ribbon(aes(ymin = conditional_mean - sqrt(conditional_var), 
                  ymax = conditional_mean + sqrt(conditional_var)), 
              alpha = 0.2, fill = "gray") +
  labs(title = "Time Series Plot with SARIMA-GARCH Conditional Mean and Conditional Variance", 
       y = "Value", 
       color = "Legend") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
par(mfrow=c(1,1))
cond_volatility <- sigma(yfit_4)
mean_model_fit <- fitted(yfit_4)
p <- plot.xts(YTs, col = "grey")
p <- addSeries(mean_model_fit, col = 'red', on = 1)
p <- addSeries(cond_volatility, col = 'blue', on = 1)
p
#plotlist<- list(p1,p2,p3)
#plot(plotlist)
#plot(x, y1, type='l')
```


#volatility forecast
```{r}
fct<-ugarchforecast(fitORspec=yfit_4,n.head=12)
#plot(fct)
plot(fitted(fct))
plot(sigma(fct))


```


```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="sGARCH",garchOrder= c(1,1)),
                  distribution.model = "snorm",)
yfit_5 <- ugarchfit(spec, YTs,solver = "hybrid")
show(yfit_5)
```

#
```{r}
#plot(yfit4,which="all")

```
#Lets try the Assymetric Power ARCH with sstd distribution

```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="apARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
yfit_6 <- ugarchfit(spec, YTs,solver = "hybrid")
show(yfit_6)
```

#Lets try the Assymetric Power ARCH with normal distribution

```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="apARCH",garchOrder= c(1,1)),
                  distribution.model = "norm",)
yfit_7 <- ugarchfit(spec, YTs,solver = "hybrid")
show(yfit_7)
```



#From the soft ware there are extensions of GARCH that were declared not suitable choices namely Threshold GARCH, Generalized Quadratic GARCH 

#our working variables will be the month.year and the port of Yuzhny so lets subset them
```{r}
gulf<-profercy.complete[,c("Month.Year","P.UREA.Arab.Gulf.FOB")]
summary(gulf)
#View(gulf)
```

#Lets write it into a ts object and visualise
#PORT OF GULF
```{r}
gulfts<-ts(gulf[,2],start=c(2004,1),end=2018,frequency = 12)
plot(gulfts,xlab="Years",ylab="price")#to plot the ts object
#abline(reg=lm(gulfts~time(gulfts)))#to insert trend line
ggtsdisplay(gulfts)
```


#Lets decompose the data for YUZ into its components
```{r}
dec<-decompose(gulfts)#to decompose the data
plot(dec)
```


#Lets test for normalility for gulf ts
```{r}
shapiro.test(gulfts)#To test for normality
```
The data is not normal since the p-value is less than 0.05(5.084e-08) so we reject the null hypothesis and conclude that the data does not follow a normal distribution

#Lets plot the histogram
```{r}
hist(gulfts)

```
The histogram confirms that indeed the data is not normal

```{r}
print(kurtosis(gulfts))
print(skewness(gulfts))
print(sd(gulfts))
library(psych)
#describe(gulfts)
```
#Lets check for outliers
```{r}
gulf<-profercy.complete[,c("Month.Year","P.UREA.Arab.Gulf.FOB")]
gulf_overall_outlier <- gulf
dim(gulf_overall_outlier)

##  Removing outliers

Q1 <- quantile(gulf_overall_outlier$P.UREA.Arab.Gulf.FOB, .25)
Q3 <- quantile(gulf_overall_outlier$P.UREA.Arab.Gulf.FOB, .75)
IQR <- IQR(gulf_overall_outlier$P.UREA.Arab.Gulf.FOB)

no_outliers <- subset(gulf_overall_outlier, gulf_overall_outlier$P.UREA.Arab.Gulf.FOB > (Q1 - 1.5*IQR) & gulf_overall_outlier$P.UREA.Arab.Gulf.FOB < (Q3 + 1.5*IQR))

dim(no_outliers)
View(no_outliers)
```
```{r}
View(gulf_overall_outlier$P.UREA.Arab.Gulf.FOB)
```


```{r}
cleangulfts<- ts(gulf_overall_outlier$P.UREA.Arab.Gulf.FOB,start=c(2004,1),end=2018,frequency = 12)
shapiro.test(cleangulfts)
skewness(cleangulfts)
```


#Transform the data by squaring
```{r}
hist(gulfts)
hist(gulfts^2)
describe(gulfts^2)
skewness(gulfts^2)
kurtosis(gulfts^2)
qqnorm(diff(gulfts^2))
qqnorm(diff(gulfts))
```
#other methods of transformation
all other methods including the squaring worsen the skewness 
```{r}
describe(sqrt(gulfts))
#head(guroot)
describe(log(gulfts))
describe(sin(gulfts))
describe(1/gulfts)
describe(gulfts^4)

```

#Lets test for stationarity
```{r}
adf.test(gulfts, k=2)#we use the adf test to check for stationarity
shapiro.test(gulfts^2)
```

From the above results the p-value(0.6035) is greater than 0.05 we therefore fail to reject our null hypothesis and conclude that the data is non stationary.

# lets take a seasonal difference
```{r}
gudiff12 = diff(gulfts,12)
gdiff12 = diff(gulfts^2,12)#  difference to make the data stationary
acf2(gdiff12, 48)
acf(gudiff12, 48)
```

The acf2 command asks for information about 48 lags. On the basis of the ACF and PACF of the 12th differences, we identified an ARIMA(2,0,0)×(0,1,1)12 model as a possibility. 

#Lets decompose to establish whether the trend is still present
```{r}
dec3<-decompose(gdiff12)
plot(dec3)
dec4<-decompose(gdiff12)
plot(dec4)
```
Trend is still present so lets take the diff to detrend the data
```{r}
gudiff1and12 = diff(gudiff12, 1) 
plot(gudiff1and12)
gdiff1and12 = diff(gdiff12, 1) 
plot(gdiff1and12)
par(mfrow=c(1,2))
acf(gdiff1and12, 48)
pacf(gdiff1and12, 48)
```
#Lets test for stationarity
```{r}
adf.test(gdiff1and12)
kpss.test(gdiff1and12)
pp.test(gdiff1and12)
```
```{r}
adf.test(gudiff1and12)
kpss.test(gudiff1and12)
pp.test(gudiff1and12)
```



#Lets select the model using non transformed auto sarima to keep away from any bias the reseacher may be having
```{r}
gugulf_auto<-auto.arima(gudiff1and12,
                 stepwise = F,
                 approximation = F,
                 seasonal = T,
                 allowdrift = F
                 )
summary(gugulf_auto)
```
The non transformed data gives a lower AIC meaning it fits the data best

```{r}
fores<-forecast(gugulf_auto, n.ahead=12)
plot(fores)
```



#Lets select the model using auto sarima to keep away from any bias the reseacher may be having using transformed data
```{r}
gulf_auto<-auto.arima(gdiff1and12,
                 stepwise = F,
                 approximation = F,
                 seasonal = T,
                 allowdrift = F
                 )
summary(gulf_auto)


```




```{r}
gupred_model<- arima(gulfts, order=c(2,1,0), seasonal = c(2,1,0))
library(forecast)
gupred_vals<- forecast(gupred_model, level=c(95), h=24)
plot(gupred_vals)
gupred_vals
```
```{r}
#Checking the validity of the forecasts
acf(gupred_vals$residuals)#To plot acf
hist(gupred_vals$residuals)#To plot histogram
Box.test(gupred_vals$residuals, lag=24, type="Ljung-Box")#To check for serial correlation

```


#Lets perform an ARCH test on the squared residuals
```{r}
ArchTest(gupred_vals$residuals)#To check for arch effects on the residuals
ArchTest(gupred_vals$residuals^2)#To check for arch effects on the squared residuals
```


#We use auto arima search so that we  can keep away any bias that may lie with the researcher
```{r}
autoplot(residuals(gulf_auto)) # plot the residuals 
plot(gulf_auto) # inspect the roots  
ggAcf(residuals(gulf_auto),lag.max=36) # inspect the ACF 
ggPacf(residuals(gulf_auto),lag.max=36) # inspect the PACF 
```
```{r}

```


#Lets fit this model
```{r}
fit1_gulf<-sarima(gulfts, 2,1,0,2,1,0,12)
fit1_gulf
```

#Lets Make Forecasts
```{r}
gpred_model<- arima(gulfts^2, order=c(2,1,0), seasonal = c(2,1,0))
library(forecast)
gpred_vals<- forecast(gpred_model, level=c(95), h=24)
plot(gpred_vals)
gpred_vals

```


```{r}
#Checking the validity of the forecasts
acf(gpred_vals$residuals)#To plot acf
hist(gpred_vals$residuals)#To plot histogram
Box.test(gpred_vals$residuals, lag=24, type="Ljung-Box")#To check for serial correlation

```
#Lets perform an ARCH test on the squared residuals
```{r}
ArchTest(gpred_vals$residuals)#To check for arch effects on the residuals
ArchTest(gpred_vals$residuals^2)#To check for arch effects on the squared residuals
```
The squared residuals pass the arch test so we proceed to fit a GARCH on the residuals

```{r}
gres<-c(gpred_vals$residuals^2)#to assign a variable name to the squared residuals
#descriptive statistics for the residuals
summary(gres)
print(kurtosis(gres))
print(skewness(gres))
print(sd(gres))
```

#Lets write the residuals into a ts object
```{r}
gts<-ts(gres)
plot(gts)
hist(gts)
acf(gts)
pacf(gts)
adf.test(gts)
plot(abs(gts))
qqnorm(gts, main = "Squared resids -QQ Plot", col = "blue")
qqline(gts)

```
```{r}
par(mfrow=c(1,2))
acf(gts)
pacf(gts)
```



#Lets specify the GARCH Model as it has ARCH effects in the residuals
`
#Lets estimate the various GARCH MODELS
#EGARCH model with student  distribution
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="eGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
gufit <- ugarchfit(spec, data=gts, solver = "hybrid")
show(gufit)
```

```{r}
gures = residuals(fit)
acf(gures)
pacf(gures)
```

#Volatility forecast
```{r}
gu<-ugarchforecast(fitORspec=fit,n.head=24)
plot(fitted(gu))
plot(sigma(gu))

```

#Lets forecast
```{r}
gu<- ugarchforecast(fit,n.ahead = 24)
gu
```

```{r}
spec1 <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="eGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
fit_1 <- ugarchfit(spec1, data=gts, solver = "hybrid")
show(fit_1)
```
#EGARCH model with skewed generalized error distribution distribution
```{r}
spec2 <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="eGARCH",garchOrder= c(1,1)),
                  distribution.model = "snorm",)
gufit2 <- ugarchfit(spec2, data=gts,solver = "hybrid")
show(gufit2)
```

#Lets forecast
```{r}
f<- ugarchforecast(fit1,n.ahead = 24)
f
```





#gjrGARCH with normal distribution
```{r}

spec3<-ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
variance.model=list(model="gjrGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd")
gufit3<-ugarchfit(spec3,data=gts,solver= "hybrid")
show(gufit3)
```

```{r}
spec4 <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="sGARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
gufit4 <- ugarchfit(spec4, data=gts,solver = "hybrid")
show(gufit4)
```
#Lets try the Assymetric Power ARCH with sstd distrib
```{r}
spec <- ugarchspec(mean.model = list(armaOrder = c(0,0), include.mean = T),
                   variance.model=list(model="apARCH",garchOrder= c(1,1)),
                  distribution.model = "sstd",)
gufit5 <- ugarchfit(spec, gts,solver = "hybrid")
show(gufit5)
```




